---
title: "Wage_exploration"
output: html_notebook
---

# Introduction

## API

We use wage data from [Canada Open Data](https://open.canada.ca/data/en/dataset/adad580f-76b0-4502-bd05-20c125de9116) website in 2012-2016 and 2019-2023.

This dataset published by the Canadian government institutions which means they are accurate, providing an overall reflection of the general state of the job market. Crawling from the website will deliver dataset to end users in JSON format.

At first, users need to apply to their own API keys (here)[]. Once you have your key, you can store it in your system environment so it is automatically used in API calls. To do so just enter `Sys.setenv(CKAN_API_KEY = "your_api_key")'`

```{r}
# Sys.setenv(CKAN_API_KEY = "522bd094-7f62-4c3b-914d-bf6f0a20dc94")
source("get_wage.R")
source("helpers.R")
source("list_occupations.R")
source("salary_range.R")
source("map.R")
source("wage_trend.R")
# source("D:/MDS/block4/Data534/project/get_wage/Data534-Project/tests/testthat.R")
```

## Accessing Wage Data

You can run `get_wage` to get datasets from Canada Open Data API. Based on what you need, you need to put a vector to scrape data, even though just one year.(e.g. `get_wage(c("2023"))`). Note that when`annual_wage_flag=0` which means the weekly payment, so multiply the `low_wage`,`high_wage`and`median_wage` columns by 40 (40 hours per week) and then 52(52 weeks per year).

You can feel safe and happy to use this function because we have do a lot data wrangling making data tidy:)

```{r}
records=get_wage(c("2012","2013","2014","2015","2016","2019","2020","2021","2022","2023"))
records
```

There are 63,748 samples in total and 7 variables to describe the data. Our main concern is occupation, median salary province and year.

This data set has 753 occupations, 14 provinces and median_wage from 19760 to 235980160.

```{r echo=FALSE}
occupation_data<-unique(tolower(records$occupation))
# occupation_data
province<-unique(records$province)
# province
summary(records$median_wage)
```

# Working with wage data

## Display the trend of wage based on provinces and occupations

### Overall View

In the first example, just need to put mandatory parameters(dataset, type=\` overall \`), the other having used default parameters to draw.

As the plot shows, there are two sub-plot used to interpret the overall trend. One is boxplot, apart from knowing the distribution of the wage,which you can also compare the number of samples in different provinces based on the size of the box; The other one is line plot which you can intuitively recognize the thread of different province.

```{r}
trend(records,"overall")
```

In the second example, just need to put mandatory parameters(dataset, `type=specific`), the other having used default parameters to draw.

As the plot shows, there are two facets describing the wage trend based on `provs=c("Ontario","British Columbia")` and `positions=c("Legislators")`. You can see the interval of wage of Legislators in Ontario is better than British Columbia. While British Columbia is more stable.

```{r}
trend(records,"specific")
```

The following examples are used to show the parameters you can change.

If you choose the `type=overall`, the only parameter you can change is `filte` which means you can decide the outliers you based on your knowledge. By the way, the default value is 100000000.

When we set `filte=200000`, it is easier to see the scenario of common people.

```{r}
trend(records,"overall",filte=200000)
```

If you choose the `type=specific`, you can also change the parameter `filte` based on your knowledge. In addition, you can also change the parameters province using `provs`and occupation using `positions`. Note both of these two parameters need to vector types.

As the plot shows, **Computer and information systems managers** is highest paid occupation, **Data entry clerks** gains lowest paid among four jobs within the range from 30000 to 60000. Data Scientists are just started occupation.

```{r}
trend(records,"specific",provs=c("Alberta","British Columbia","Ontario","Quebec"),positions=c("Data entry clerks","Database analysts and data administrators","Data scientists","Computer and information systems managers"))
```

```{r}
# install.packages("covr")
library(covr)
report()


```

```{r}
map(records,"Legislators")
```

```{r}
setwd("D:/MDS/block4/Data534/project/get_wage/Data534-Project/R")
```

```{r}
library(covr)
cov <- package_coverage()

```
